---
title: "Project 3"
author: "Yuhan Xiao"
output: html_document
date: "2023-10-22"
---

## Background 
```{r message = FALSE, warning = FALSE}
library(here)
library(tidyverse)
library(lubridate)
library(ggplot2)
```

#### Load data
```{r message = FALSE}
rds_files <- c("b_lyrics.RDS", "ts_lyrics.RDS", "sales.RDS")
## Check whether we have all 3 files
if (any(!file.exists(here("data", rds_files)))) {
    ## If we don't, then download the data
    b_lyrics <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/beyonce_lyrics.csv")
    ts_lyrics <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/taylor_swift_lyrics.csv")
    sales <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/sales.csv")

    ## Then save the data objects to RDS files
    saveRDS(b_lyrics, file = here("data", "b_lyrics.RDS"))
    saveRDS(ts_lyrics, file = here("data", "ts_lyrics.RDS"))
    saveRDS(sales, file = here("data", "sales.RDS"))
}

# load the datasets
b_lyrics <- readRDS(here("data", "b_lyrics.RDS"))
ts_lyrics <- readRDS(here("data", "ts_lyrics.RDS"))
sales <- readRDS(here("data", "sales.RDS"))
```

## Part 1: Explore album sales
In this section, the goal is to explore the sales of studio albums from Beyoncé and Taylor Swift.

**Notes**

* In each of the subsections below that ask you to create a plot, you must create a title, subtitle, x-axis label, and y-axis label with units where applicable. For example, if your axis says “sales” as an axis label, change it to “sales (in millions)”.

#### Part 1A
In this section, we will do some data wrangling.

1. Use `lubridate` to create a column called `released` that is a `Date` class. However, to be able to do this, you first need to use `stringr` to search for pattern that matches things like this “(US)[51]” in a string like this “September 1, 2006 (US)[51]” and removes them. (**Note**: to get full credit, you must create the regular expression).
2. Use `forcats` to create a factor called country (**Note**: you may need to collapse some factor levels).
3. Transform the `sales` into a unit that is album sales in millions of dollars.
4. Keep only album sales from the UK, the US or the World.
5. Auto print your final wrangled tibble data frame.

```{r}
pattern = "\\s\\(\\w+\\)\\[\\d+\\]"
sales_wrangled <- sales %>%
  mutate(released = str_replace(released, pattern, "")) %>%
  mutate(released = mdy(released)) %>%
  mutate(country = as.factor(country)) %>%
  mutate(country = fct_collapse(country, "FRA" = c("FR", "FRA"),
         "WW" = c("World", "WW"))) %>%
  mutate(sales = sales/1000000) %>%
  filter(country %in% c("UK", "US", "WW"))

sales_wrangled
```

#### Part 1B
In this section, we will do some more data wrangling followed by summarization using wrangled data from Part 1A.

1. Keep only album sales from the US.
2. Create a new column called `years_since_release` corresponding to the number of years since the release of each album from Beyoncé and Taylor Swift. This should be a whole number and you should round down to “14” if you get a non-whole number like “14.12” years. (**Hint**: you may find the `interval()` function from `lubridate` helpful here, but this not the only way to do this.)
3. Calculate the most recent, oldest, and the median years since albums were released for both Beyoncé and Taylor Swift.
```{r}
sales_summ <- sales_wrangled %>%
  select(country, released, artist) %>%
  filter(country == "US") %>%
  mutate(years_since_release = floor(as.duration(interval(released, today())) / dyears(1))) %>%
  group_by(artist) %>%
  summarise(
    most_recent_years = min(years_since_release),
    oldest_years = max(years_since_release),
    median_years = median(years_since_release)
  )

sales_summ
```

#### Part 1C
Using the wrangled data from Part 1A:

1. Calculate the total album sales for each `artist` and for each `country` (only sales from the UK, US, and World).
2. Using the total album sales, create a percent stacked barchart using `ggplot2` of the percentage of sales of studio albums (in millions) along the y-axis for the two artists along the x-axis colored by the `country`.

```{r message = FALSE}
sales_total <- sales_wrangled %>%
  select(artist, country, sales) %>%
  filter(country %in% c("UK", "US", "WW")) %>%
  group_by(artist, country) %>%
  summarise(
    total_sales = sum(sales)
  )

sales_total %>% ggplot(aes(fill=country, y=total_sales, x=artist)) +
  geom_bar(position="fill", stat="identity") +
  scale_fill_manual(values = c("#FAAB18", "#4CAF50", "#1380A1")) +
  labs(title = "Total Album Sales Percent for Each Artist by Country", subtitle = str_wrap("Beyoncé has a higher percent of album sales from the world than Taylor Swift", 50), caption = "Yuhan Xiao", x = "artist", y = "sales (in percent)", color = "country") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
        axis.text.x = element_text(angle = 0, hjust = 1),
        plot.background = element_rect(fill = "white"),
        text = element_text(family = "Times"))
  
```
