---
title: "Project 3"
author: "Yuhan Xiao"
output: html_document
date: "2023-10-22"
---

## Background 
```{r bg, message = FALSE, warning = FALSE}
library(here)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(tidytext)
library(wordcloud)
```

#### Load data
```{r load, message = FALSE}
rds_files <- c("b_lyrics.RDS", "ts_lyrics.RDS", "sales.RDS")
## Check whether we have all 3 files
if (any(!file.exists(here("data", rds_files)))) {
    ## If we don't, then download the data
    b_lyrics <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/beyonce_lyrics.csv")
    ts_lyrics <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/taylor_swift_lyrics.csv")
    sales <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/sales.csv")

    ## Then save the data objects to RDS files
    saveRDS(b_lyrics, file = here("data", "b_lyrics.RDS"))
    saveRDS(ts_lyrics, file = here("data", "ts_lyrics.RDS"))
    saveRDS(sales, file = here("data", "sales.RDS"))
}

# load the datasets
b_lyrics <- readRDS(here("data", "b_lyrics.RDS"))
ts_lyrics <- readRDS(here("data", "ts_lyrics.RDS"))
sales <- readRDS(here("data", "sales.RDS"))
```

## Part 1: Explore album sales
In this section, the goal is to explore the sales of studio albums from Beyoncé and Taylor Swift.

**Notes**

* In each of the subsections below that ask you to create a plot, you must create a title, subtitle, x-axis label, and y-axis label with units where applicable. For example, if your axis says “sales” as an axis label, change it to “sales (in millions)”.

#### Part 1A
In this section, we will do some data wrangling.

1. Use `lubridate` to create a column called `released` that is a `Date` class. However, to be able to do this, you first need to use `stringr` to search for pattern that matches things like this “(US)[51]” in a string like this “September 1, 2006 (US)[51]” and removes them. (**Note**: to get full credit, you must create the regular expression).
2. Use `forcats` to create a factor called country (**Note**: you may need to collapse some factor levels).
3. Transform the `sales` into a unit that is album sales in millions of dollars.
4. Keep only album sales from the UK, the US or the World.
5. Auto print your final wrangled tibble data frame.

```{r 1a}
pattern = "\\s\\(\\w+\\)\\[\\d+\\]"
sales_wrangled <- sales %>%
  mutate(released = str_replace(released, pattern, "")) %>%
  mutate(released = mdy(released)) %>%
  mutate(country = as.factor(country)) %>%
  mutate(country = fct_collapse(country, "FRA" = c("FR", "FRA"),
         "WW" = c("World", "WW"))) %>%
  mutate(sales = sales/1000000) %>%
  filter(country %in% c("UK", "US", "WW"))

sales_wrangled
```

#### Part 1B
In this section, we will do some more data wrangling followed by summarization using wrangled data from Part 1A.

1. Keep only album sales from the US.
2. Create a new column called `years_since_release` corresponding to the number of years since the release of each album from Beyoncé and Taylor Swift. This should be a whole number and you should round down to “14” if you get a non-whole number like “14.12” years. (**Hint**: you may find the `interval()` function from `lubridate` helpful here, but this not the only way to do this.)
3. Calculate the most recent, oldest, and the median years since albums were released for both Beyoncé and Taylor Swift.
```{r 1b}
sales_summ <- sales_wrangled %>%
  select(country, released, artist) %>%
  filter(country == "US") %>%
  mutate(years_since_release = floor(as.duration(interval(released, today())) / dyears(1))) %>%
  group_by(artist) %>%
  summarise(
    most_recent_years = min(years_since_release),
    oldest_years = max(years_since_release),
    median_years = median(years_since_release)
  )

sales_summ
```

#### Part 1C
Using the wrangled data from Part 1A:

1. Calculate the total album sales for each `artist` and for each `country` (only sales from the UK, US, and World).
2. Using the total album sales, create a percent stacked barchart using `ggplot2` of the percentage of sales of studio albums (in millions) along the y-axis for the two artists along the x-axis colored by the `country`.

```{r 1c, message = FALSE}
sales_total <- sales_wrangled %>%
  select(artist, country, sales) %>%
  filter(country %in% c("UK", "US", "WW")) %>%
  group_by(artist, country) %>%
  summarise(
    total_sales = sum(sales)
  )

sales_total %>% ggplot(aes(fill=country, y=total_sales, x=artist)) +
  geom_bar(position="fill", stat="identity") +
  scale_fill_manual(values = c("#FAAB18", "#4CAF50", "#1380A1")) +
  labs(title = "Total Album Sales Percent for Each Artist by Country", subtitle = str_wrap("Beyoncé has a higher percent of album sales from the world than Taylor Swift", 50), caption = "Yuhan Xiao", x = "artist", y = "sales (in percent)", fill = "country") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
        axis.text.x = element_text(angle = 0, hjust = 1),
        plot.background = element_rect(fill = "white"),
        text = element_text(family = "Times"))
  
```

#### Part 1D
Using the wrangled data from Part 1A, use `ggplot2` to create a bar plot for the sales of studio albums (in millions) along the x-axis for each of the album titles along the y-axis.

**Note**:

* You only need to consider the global World sales (you can ignore US and UK sales for this part).
* The title of the album must be clearly readable along the y-axis.
* Each bar should be colored by which artist made that album.
* The bars should be ordered from albums with the most sales (top) to the least sales (bottom) (**Note**: you must use functions from `forcats` for this step).

```{r 1d}
sales_wrangled %>%
  select(sales, artist, title, country) %>%
  filter(country == "WW") %>%
  ggplot(aes(sales, fct_reorder(.f = title, .x = sales), fill = fct_reorder2(artist, sales, title))) +
    geom_bar(stat = "identity") + 
    scale_fill_manual(values = c("#FAAB18", "#1380A1")) +
    labs(title = "Total Album Sales (in Millions) for Each Artist by Album Title", subtitle = str_wrap("Taylor Swift has more album sales than Beyoncé overall", 60), caption = "Yuhan Xiao", x = "sales (in millions)", y = "album title", fill = "artist") +
    theme_minimal(base_size = 14) +
    theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
          plot.subtitle = element_text(hjust = 0.5),
          axis.text.x = element_text(angle = 0, hjust = 1),
          plot.background = element_rect(fill = "white"),
          text = element_text(family = "Times"))
```

#### Part 1E
Using the wrangled data from Part 1A, use `ggplot2` to create a scatter plot of sales of studio albums (in millions) along the y-axis by the released date for each album along the x-axis.

**Note**:

* The points should be colored by the artist.
* There should be three scatter plots (one for UK, US and world sales) faceted by rows.

```{r 1e}
sales_wrangled %>%
  select(sales, released, artist, country) %>%
  ggplot(aes(released, sales, color = fct_reorder2(artist, sales, released))) +
    facet_wrap(~country, nrow=2, scales="fixed") +
    geom_point() +
    scale_color_manual(values = c("#FAAB18", "#1380A1")) +
    labs(title = str_wrap("Total Album Sales (in Millions) Over Album Released Date for Each Artist by Country", 70), subtitle = str_wrap("Taylor Swift has more album sales than Beyoncé overall. Beyoncé released a few albums before Taylor Swift released hers", 70), caption = "Yuhan Xiao", x = "released date (year)", y = "sales (in millions)", color = "artist") +
    theme_minimal(base_size = 14) +
    theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
          plot.subtitle = element_text(hjust = 0.5),
          axis.text.x = element_text(angle = 0, hjust = 1),
          plot.background = element_rect(fill = "white"),
          text = element_text(family = "Times"),
          legend.position = c(0.8, 0.15))
  
```

## Part 2: Exploring sentiment of lyrics
In Part 2, we will explore the lyrics in the `b_lyrics` and `ts_lyrics` datasets.

#### Part 2A
Using `ts_lyrics`, create a new column called `line` with one line containing the character string for each line of Taylor Swift’s songs.

* How many lines in Taylor Swift’s lyrics contain the word “hello”? For full credit, show all the rows in `ts_lyrics` that have “hello” in the `line` column and report how many rows there are in total.
* How many lines in Taylor Swift’s lyrics contain the word “goodbye”? For full credit, show all the rows in `ts_lyrics` that have “goodbye” in the `line` column and report how many rows there are in total.

```{r 2a}
ts_lyrics_line <- ts_lyrics %>%
  unnest_tokens(
    output = line,
    input = Lyrics,
    token = "lines"
  ) 

pattern = "hello"
ts_lyrics_line %>% 
  select(Artist, line) %>%
  filter(grepl(pattern, line)) %>%
  mutate(nrow = n())

pattern = "goodbye"
ts_lyrics_line %>% 
  select(Artist, line) %>%
  filter(grepl(pattern, line)) %>%
  mutate(nrow = n())
```

#### Part 2B
Repeat the same analysis for b_lyrics as described in Part 2A.

```{r 2b}
pattern = "hello"
b_lyrics %>% 
  select(artist_name, line) %>%
  filter(grepl(pattern, line)) %>%
  mutate(nrow = n())

pattern = "goodbye"
b_lyrics %>% 
  select(artist_name, line) %>%
  filter(grepl(pattern, line)) %>%
  mutate(nrow = n())
```

#### Part 2C
Using the `b_lyrics` dataset,

1. Tokenize each lyrical line by words.
2. Remove the “stopwords”.
3. Calculate the total number for each word in the lyrics.
4. Using the “bing” sentiment lexicon, add a column to the summarized data frame adding the “bing” sentiment lexicon.
5. Sort the rows from most frequent to least frequent words.
6. Only keep the top 25 most frequent words.
7. Auto print the wrangled tibble data frame.
8. Use `ggplot2` to create a bar plot with the top words on the y-axis and the frequency of each word on the x-axis. Color each bar by the sentiment of each word from the “bing” sentiment lexicon. Bars should be ordered from most frequent on the top to least frequent on the bottom of the plot.
9. Create a word cloud of the top 25 most frequent words.

```{r 2c, message = FALSE}
b_top_25_freq_word <- b_lyrics %>%
  unnest_tokens(
    output = word,
    input = line
  ) %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE) %>%
  inner_join(get_sentiments("bing")) %>%
  head(25)

b_top_25_freq_word

b_top_25_freq_word %>%
  ggplot(aes(n, fct_reorder(.f = word, .x = n), fill = fct_reorder2(sentiment, word, n))) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("#FAAB18", "#1380A1")) +
  labs(title = str_wrap("The Top 25 Most Frequent Words in Beyoncé's Lyrics", 70), subtitle = str_wrap("More words are positive rather than negative.", 70), caption = "Yuhan Xiao", x = "frequency", y = "word", fill = "sentiment") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 0, hjust = 1),
        plot.background = element_rect(fill = "white"),
        text = element_text(family = "Times"))

b_top_25_freq_word %>%
  with(wordcloud(word, n))
```

#### Part 2D
Repeat the same analysis as above in Part 2C, but for `ts_lyrics`.

```{r 2d, message = FALSE}
ts_top_25_freq_word <- ts_lyrics_line %>%
  unnest_tokens(
    output = word,
    input = line
  ) %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE) %>%
  inner_join(get_sentiments("bing")) %>%
  head(25)

ts_top_25_freq_word

ts_top_25_freq_word %>%
  ggplot(aes(n, fct_reorder(.f = word, .x = n), fill = fct_reorder2(sentiment, word, n))) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("#FAAB18", "#1380A1")) +
  labs(title = str_wrap("The Top 25 Most Frequent Words in Taylor Swift's Lyrics", 70), subtitle = str_wrap("The word 'love' is the most frequently used one. Other words are more negative than positive.", 70), caption = "Yuhan Xiao", x = "frequency", y = "word", fill = "sentiment") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 0, hjust = 1),
        plot.background = element_rect(fill = "white"),
        text = element_text(family = "Times"))

ts_top_25_freq_word %>%
  with(wordcloud(word, n))
```

#### Part 2E
Using the `ts_lyrics` dataset,

1. Tokenize each lyrical line by words.
2. Remove the “stopwords”.
3. Calculate the total number for each word in the lyrics **for each Album**.
4. Using the “afinn” sentiment lexicon, add a column to the summarized data frame adding the “afinn” sentiment lexicon.
5. Calculate the average sentiment score **for each Album**.
6. Auto print the wrangled tibble data frame.
7. Join the wrangled data frame from Part 1A (album sales in millions) with the wrangled data frame from #6 above (average sentiment score for each album).
8. Using `ggplot2`, create a scatter plot of the average sentiment score for each album (y-axis) and the album release data along the x-axis. Make the size of each point the album sales in millions.
9. Add a horizontal line at y-intercept=0.
10. Write 2-3 sentences interpreting the plot answering the question “How has the sentiment of Taylor Swift’s albums have changed over time?”. Add a title, subtitle, and useful axis labels.

```{r 2e, message = FALSE}
ts_album_ss <- ts_lyrics_line %>%
  unnest_tokens(
    output = word,
    input = line
  ) %>%
  anti_join(stop_words) %>%
  group_by(Album) %>%
  count(word, sort = TRUE) %>%
  inner_join(get_sentiments("afinn")) %>%
  summarize(
    avg_sentiment = mean(value)
  ) %>%
  mutate(Album = tolower(Album))

ts_album_ss

sales_wrangled %>%
  rename(Album = title) %>%
  mutate(Album = tolower(Album)) %>%
  inner_join(ts_album_ss) %>%
  ggplot(aes(released, avg_sentiment, size = sales)) +
    geom_point(color = "#FAAB18") +
    geom_hline(yintercept = 0, linetype = "dashed", color = "#1380A1") +
    labs(title = str_wrap("The Average Sentiment Score of Taylor Swift's Album Over Released Year", 60), subtitle = str_wrap("Taylor Swift's albums overall have a negative sentiment. The album '1989' released in 2014 is the most negative one. The two more recent albums 'Reputation' and 'Lover' are more positive though they are still less positive than her early albums released before '1989'.", 70), caption = "Yuhan Xiao", x = "released date (year)", y = "average sentiment score", size = "sales (in millions)") +
    theme_minimal(base_size = 14) +
    theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15),
          plot.subtitle = element_text(hjust = 0.5),
          axis.text.x = element_text(angle = 0, hjust = 1),
          plot.background = element_rect(fill = "white"),
          text = element_text(family = "Times"))
```